---
title: "Assignment 3"
author: "Brashon Ford"
date: August 20th
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Read in data
myDataLocation <- "c:\pathtofolderlocation"
setwd(myDataLocation)
dat <- read.csv(file = "cancer_data.csv", header = TRUE)

# Load and install required packages
library(mosaic)
#install.packages("effsize")
library(effsize)
library(lsr)
library(supernova)
```

## Question 1
1.	Descriptive statistics.
  a.	Quantitative Variables: perimeter_mean and texture_mean
    i.	Create a histogram for each variable
    ii.	Obtain the five number summaries for each variable
  b.	Categorical Variable: diagnosis
   i.	Create a boxplot with perimeter_mean as the explanatory variable
    ii.	Obtain frequencies of the two categories.
    iii.	Obtain means and standard deviations of perimeter_mean, aggregated by diagnosis
```{r}
#1a i and ii. 
hist(cancer_data$perimeter_mean , data = cancer_data)
hist(cancer_data$texture_mean , data = cancer_data)
summary(cancer_data$perimeter_mean)
summary(cancer_data$perimeter_mean)
boxplot(perimeter_mean ~ diagnosis , data = cancer_data,
        xlab = "Diagnosis" , ylab = "Perimeter Mean",
        main = "Boxplot of Perimeter Mean by Diagnosis")
frequency(cancer_data$perimeter_mean)
frequency(cancer_data$diagnosis) 

#1 b
agg.result <- aggregate(perimeter_mean ~ diagnosis , data = cancer_data,
          FUN = function(x) c(mean = mean(x), sd(x))) 

print(agg.result) 
```

## Question 2

2.	Correlation between perimeter_mean and texture_mean
  a.	Obtain the correlation.
  b.	Interpret the correlation (reference Assignment One for examples).
  c.	What is the 95% confidence interval? Use cor.test().
  d.	What is the 75% confidence interval? Use cor.test().
  e.	Are you confident that the true population correlation could be zero? Why or why not?


```{r}
cor(cancer_data$perimeter_mean , cancer_data$texture_mean)
cor.test(cancer_data$perimeter_mean , cancer_data$texture_mean , conf.level = 0.75)
cor.test(cancer_data$perimeter_mean , cancer_data$texture_mean , conf.level = 0.95)
#Q2 
#a 
0.334
#b.The correlation is pretty close to 0, which can suggest that there is a weak relationship between the perimeter mean and diagnosis. It is a positive correlation. 
#c.  0.2597318 0.4059340
#d 0.29-0.37
#e I believe it would be true that the population correlation could be zero just based on the p-value that we received. It is a very small number, not too close to 0.05
```

## Question 3

3.	Fit an empty model for perimeter_mean,
  a.	What are the parameter estimates? Interpret them.
  b.	What is the PRE? Interpret it. 
  c.	Provide a 95% confidence interval around your parameter estimate. How did you obtain it? Interpret it.

```{r}
empty_model <- lm(perimeter_mean ~ 1 , data = cancer_data)
summary(empty_model)

#a
#The parameter estimates just include the intercept which is 92.05
#b
#the PRE for this model is 0, there is no value for the PRE in the anova table
#c
#	             2.5 %   97.5 %
#	(Intercept) 90.04806 94.04511
#We can see we only get one group of intervals; primarily because there is only one parameter. We can see that the lower bound is 90.04 and the upper bound is 94.04.

```
## Question 4

4.	Fit a linear model with perimeter_mean as the outcome and diagnosis as the explanatory variable.
  a.	What are the parameter estimates? Interpret them.
  b.	What are the 95% confidence intervals around your slope? Describe the method you used to obtain them.
  c.	What are the 99% confidence intervals around your slope? Describe the method you used to obtain them.
  d.	What is the effect size, expressed as Cohen’s d?
    i.	Interpret the Cohen’s d from the model.
    ii.	What is the 99% confidence interval around Cohen’s d?
    iii.	Interpret the confidence interval for Cohen’s d. What do the units represent?
  e.	Do you believe there is a relationship between perimeter_mean and diagnosis? Why or why not?
  f.	How does the information you received from the confidence interval relate to the hypothesis that this sample represents (is drawn from) a population with these parameter estimates?

```{r}
lm(perimeter_mean ~ diagnosis , data = cancer_data)
cor.test(cancer_data$perimeter_mean , cancer_data$texture_mean , conf.level = 0.95)
cor.test(cancer_data$perimeter_mean , cancer_data$texture_mean, conf.level = 0.99) 
effect_size <- effsize::cohen.d(cancer_data$perimeter_mean , cancer_data$diagnosis)
print(effect_size)

#b and c The method I used to obtain the confidence intervals was by using the cor.test function and then making sure to put that I needed 95% intervals. 95% percent confidence interval:     0.2597318 0.4059340. The method that I used to obtain the data is the confint() function. 99 percent confidence interval:  0.2354118 0.4273442


#d The effect size is -2.28 the cohen’s d is pretty large for this model; that can explain that the two groups are in completely different directions.
#e 

#f - Skip
```
## Question 5

5.	Fit a linear model with perimeter_mean as the outcome and texture_mean as the explanatory variable.
  a.	What are the parameter estimates? Interpret them.
  b.	95% confidence intervals around your slope.
    i.	Construct the confidence intervals using bootstrap with 1000 samples. 
      1.	What are the units of the confidence intervals?
      2.	How would you interpret the confidence interval?
    ii.	Construct the confidence intervals using the model. 
      1.	What are the units of the confidence intervals?
      2.	How would you interpret the confidence interval?
    iii.	Are you confident that the true population slope parameter is not 0? Why or why not?
  c.	75% confidence intervals around your slope.
    i.	Construct the confidence intervals using bootstrap with 1000 samples.
    ii.	Construct the confidence intervals using the model.
    iii.	Are the confidence intervals from these two methods similar? If they are similar, why is this so? If they are not, why not?
  d.	What is the effect size, expressed as PRE? Interpret the PRE of the model.
  e.	Do you believe there is a relationship between texture_mean and perimeter_mean? Why or why not?

```{r paged.print=FALSE}
linear_model <- lm(perimeter_mean ~ texture_mean , data = cancer_data)
summary(linear_model)

data <- data.frame(texture_mean, perimeter_mean)  

slope_function <- function(data, indices) {
  sampled_data <- data[indices, 1000 ]
  model <- lm(perimeter_mean ~ texture_mean, data = sampled_data)
  return(coef(model)[2])  
}

boot_result <- boot(data, slope_function, R = 1000) 
conf_interval <- quantile(boot_result$t, c(0.025, 0.975))
print("95% Confidence Interval for the Slope:")
print(conf_interval)



data <- data.frame(texture_mean, perimeter_mean)  


slope_function <- function(data, indices) {
  sampled_data <- data[indices, ]
  model <- lm(perimeter_mean ~ texture_mean, data = sampled_data)
  return(coef(model)[2]) 
}


boot_result <- boot(data, slope_function, R = 1000)  

conf_interval <- quantile(boot_result$t, c(0.025, 0.75))


print("75% Confidence Interval for the Slope:")
print(conf_interval)



linear.model.2 <- lm(perimeter_mean ~ texture_mean , data = data)
confint(linear.model.2 , level = 0.75)


model_summary <- summary(linear.model.2)
pre <- model_summary$r.squared
print(pre)

#Q5 
#a For the Intercept I got 55.631 and for the texture mean increment its 1.889.

#bi: The confidence intervals are 1.88-1.88; the units of the confidence intervals would be meters per unit.
#bii: the confidence intervals show where the the range of values are within the 95%.
#biii: I believe that the true population is not 0 because the 95% confidence interval does not include 0 at all.



#c: The confidence intervals for 75% include: 1.45--2.03(This is for the bootstrap model)
#cii: The confidence intervals for the model include 1.63--2.14.
#ciii: The two confidence intervals are kind of similar but they are a bit off. The model CI's are a bit moved to the right from the bootstrapped model. They are different just because the bootstrapped model has more samples compared to the original model.


#d The effect size, in PRE, is 0.1121223
#di: I dont believe that the perimeter_mean and texture_mean have a relationship due to the PRE being pretty small. 


```
## Question 6

6.	Comparing the models in Q4 and Q5, which do you believe predicts perimeter_mean better? Why or why not? Justify your reasoning.

```{r paged.print=FALSE}
```
I think that texture_mean predicts perimeter_mean a bit better. Just based on the data we got.
